---
title: "simpleRegress"
author: "Shawn Olson"
date: "10/21/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Model WEIGHT ~ AGE  this is an example  
```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```

Best to worst
1.Neck-> ARsqrd:0.87, RST:43.91
2.Age -> ARsqrd:0.55, RST:81.47
The agemodel graph shows its points to be far away from the line, while the simp graph has plots closer to it.


```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```
Best to worst
1. Neck -> ARsqrd:0.87. RST:43.91
2. Head length-> ARsqrd:0.69, RST:67.81
3.Age - ARsrd: 0.55, RST:81.47
Although the headlenmodel graph has scattered out plots it is better than the graph of agemodel. Comparing the Adjusted R-squared being higher at 0.69 and the agemodel Adjusted R-squared is 0.55, shows it being better than it but not high enough for simp’s Adjusted R-squared of 0.87.

## Model WEIGHT ~ HEADWTH

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH, data=bears)
plot (WEIGHT~HEADWTH, data=bears)
abline (headwthmodel)
summary.lm(headwthmodel)
```
Best to worst
1. Neck -> ARsqrd:0.87, RST:43.91
2. Head length-> ARsqrd:0.69, RST:67.81
3. Head width - ARsard:0.60, RST:76.41
4.Age -> ARsrd: 0.55, RST:81.47
3 Publish -
headwthmodel would have a high Residual standard error of 76.41, being lower than agemodel’s 81.47 but not low enogh for headlenmodel’s
67.81 Residual standard error.

## Model WEIGHT ~ LENGTH

```{r}
lengthmodel <- Im(WEIGHT~LENGTH, data=bears)
plot(WEIGHT~LENGTH, data=bears)
abline (lengthmodel)
summary.lm(lengthmodel)
