---
title: "simpleRegress"
author: "Shawn Olson"
date: "10/21/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Model WEIGHT ~ AGE  this is an example  
```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```

Best to worst
1.Neck-> ARsqrd:0.87, RST:43.91
2.Age -> ARsqrd:0.55, RST:81.47
The agemodel graph shows its points to be far away from the line, while the simp graph has plots closer to it.


```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)





